Analytic Light Curves in Reflected Light: \\% Phase Curves, Occultations, and Non-Lambertian Scattering

### 2020-05-11 tex/ms.tex
new fig

%
\section{Introduction}
\label{sec:intro}
%
\xxx{A few paragraphs of introduction here. \citet{Russell1906}.}

@@

Our implementation of the scattering model extends just as easily
to occultations and to cases where the surface does not have uniform albedo.
Figure~\ref{fig:io_europa} shows a visible-light observation of the
occultation of Io by Europa on 04 Dec 2009 taken by the PHEMU09
campaign \citep{Arlot2014}. The trajectory of Europa relative to Io,
computed from the JPL Horizons databse \xxx{(cite)}
is shown at the top. We fit to this data a \starry occultation model
with the scattering law discussed above.
For simplicity, we set the surface map equal too
an $l=15$ spherical harmonic expansion of the Galileo global color mosaic
of Io \citep{Becker2005} and use ephemerides from the JPL Horizons
database, allowing for a small static $x-y$ offset between Europa and Io as
in \citep{Arlot2014} due to the uncertainty in the database. In total,
we fit for five parameters: the two Cartesian offset terms,
the flux contribution from Europa, the average
albedo of Io, and the average surface roughness of Io, $\sigma$. The model
is displayed in blue and closely matches the data.
%
Note, however, that this is meant simply as a demonstration of the \starry
algorithm, as our model for the surface is approximate at best, given
differences in the wavelength band between the Galileo observations and
those of the PHEMU09 campaign, changes in the albedo of Io since the Galileo
measurements, and the fact that the orientation and extent of any shadows
due to volcanoes on the surface are likely different between the Galileo
and PHEMU09 observations. Furthermore,
proper modeling would entail the
joint analysis of all light curves of Io taken in a given season, for which
we can afford to simultaneously fit for the surface map without risk of
overfitting (Bartoli\'c, Luger, et al., in prep).

@@

\subsection{Information theory}
\label{sec:information}

\xxx{A few notes here on how constraining reflected light curves can be.}

@@

>>>>

%
\section{Introduction}
\label{sec:intro}
%
\xxx{A few paragraphs of introduction here.
    Papers to cite: \citet{Russell1906}, \citet{Cowan2009},
    \citet{Kawahara2010}, \citet{Kawahara2011}, \citet{Fujii2012}}

@@

Our implementation of the scattering model extends just as easily
to occultations and to cases where the surface does not have uniform albedo.
Figure~\ref{fig:io_europa} shows a visible-light observation of the
occultation of Io by Europa on 04 Dec 2009 taken by the PHEMU09
campaign \citep{Arlot2014}. The trajectory of Europa relative to Io,
computed from the JPL Horizons databse \xxx{(cite)}
is shown at the top. We fit to this data a \starry occultation model
with the scattering law discussed above.
For simplicity, we set the surface map equal too
an $l=15$ spherical harmonic expansion of the Galileo global color mosaic
of Io \citep{Becker2005} and use ephemerides from the JPL Horizons
database, allowing for a small static $x-y$ offset between Europa and Io as
in \citep{Arlot2014} due to the uncertainty in the database. In total,
we fit for five parameters: the two Cartesian offset terms,
the flux contribution from Europa, the average
albedo of Io, and the average surface roughness of Io, $\sigma$. The model
is displayed in blue and closely matches the data.
%
Note, however, that this is meant simply as a demonstration of the \starry
algorithm, as our model for the surface is approximate at best, given
differences in the wavelength band between the Galileo observations and
those of the PHEMU09 campaign, changes in the albedo of Io since the Galileo
measurements, and the fact that the orientation and extent of any shadows
due to volcanoes on the surface are likely different between the Galileo
and PHEMU09 observations. Furthermore,
proper modeling would entail the
joint analysis of all light curves of Io taken in a given season, for which
we can afford to simultaneously fit for the surface map without risk of
overfitting (Bartoli\'c, Luger et al., in prep).

@@

\subsection{Linearity}
\label{sec:linearity}

\xxx{The starry problem is linear. We compute fluxes but can easily
    compute design matrices instead.}

\subsection{The information content of reflected light curves}
\label{sec:information}

One of the fundamental difficulties with the mapping problem is that
the process of inferring a two-dimensional map from a light curve
is almost always ill-posed. This has been known since at least
the turn of the last century, when \citet{Russell1906} discussed
how, because of symmetry,
all odd harmonics above $l = 1$ are in the null space for thermal
phase curves of spherical bodies, meaning those terms do not contribute
at all to the disk-integrated flux. As discussed in \citet{Luger2019},
the problem is even more ill-posed than that: for any even
degree $l > 0$, there are $2l + 1$ modes on the \emph{surface}
(one for each value of $m$), but only $2$ Fourier modes in the
\emph{light curve} (i.e., a sine and a cosine). Thus, for every
mode that can be constrained from the light curve, there are far
more modes that cannot, a problem that only gets worse as $l$
increases.

@@

The left panel of Figure~\ref{fig:inference} shows this issue in practice.
We generate a mock thermal light curve (center left) from an $l = 20$
expansion of the Earth (top) with 1,000 points over the course of one year
with an exquisite photometric precision of 1 ppm. The Earth is given
an obliquity of $23.5^\circ$ on the plane of the sky but is viewed
along the ecliptic, rotating edge-on with an inclination of $90^\circ$.
The data are shown in black, and in blue is our least-squares fit,
in which we assume a prior variance of $10^{-3}$ for all spherical
harmonic coefficients (and zero covariance). The corresponding surface
map is shown at the bottom. As expected, this looks nothing like the
true map of the Earth. For a body seen rotating edge-on, the information
content of the light curve is strictly longitudinal. While the
inferred map captures the average brightness of the Earth at each
longitude fairly well, it is missing all latitudinal information.
This is independent of the signal-to-noise or the cadence of the
dataset---it is a fundamental limitation of phase curves in thermal light.

The same is not true for the case of reflected light phase curves. In
the right panel of Figure~\ref{fig:inference} we show the same
mock light curve and perform the same inference step,
but this time for observations in reflected light; the light curve
is similar to that in Figure~\ref{fig:earthphase}. Because of the
presence of a day-night terminator beyond which features on the surface
contribute zero flux, none of the symmetry arguments above apply. In
particular, the facts that (1) the Earth is seen at different phases and
(2) the terminator is inclined relative to the rotational axis mean
that the region of the surface contributing to the phase curve is
always changing, resulting in a complex light curve that encodes
signficicantly more information than its thermal counterpart.
%
The result is an inferred map that is largely faithful to the
true map (bottom). While the continental outlines are somewhat fuzzy and some
artifacts are present at high latitudes, it is clear that the
inference problem is much less ill-posed in this case.

Unlike thermal light curves, reflected light phase curves have the
potential to robustly constrain two-dimensional maps of exoplanets.
This result is
not new, and has been discussed at length in the literature
\citep[e.g.,][]{Fujii2012,Berdyugina2017,Luger2019b,Aizawa2020,Kawahara2020}.
In particular, \citet{Kawahara2010} demonstrated the uniqueness of their
inferred map from mock reflected light curves of the Earth. As we argued
above, for specific geometrical configurations, the mapping problem
in reflected light can actually be well-posed, meaning it has no
null space up to a certain degree $l$. Reflected light phase curves of
terrestrial planets with JWST and future direct imaging missions thus
have the potential to reveal detailed information about their surfaces.

\begin{figure}[p!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/inference_occ.pdf}
        \oscaption{inference_occ}{%
            Similar to Figure~\ref{fig:inference}, but for 500 high
            signal to noise observations taken
            over ten days near quadrature.
            At the top we show the thermal and reflected
            phase curves and the corresponding inferred maps.
            At the bottom, we show the same light curves but this
            time including 7 equatorial occultations by
            a (very) short period moon one-quarter the size of the planet.
            %
            The presence of the occultations significantly increases the
            fidelity of the recovered maps near the equator (the path
            of the occultor). While in the thermal case there is significant
            overfitting at high latitudes, in the reflected case the map
            accurately recovers features across the entire planet.
            \label{fig:inference_occ}
        }
    \end{centering}
\end{figure}

Nevertheless, at finite signal to noise and for limited observation
duration or cadence, there may still be significant degeneracies in
the reflected phase curve problem. In \citet{Luger2019} we argued that
occultations can be used to break many of these degeneracies, since they
directly probe the surface at scales inaccessible to phase curves.
The same is true in reflected light. Figure~\ref{fig:inference_occ}
demonstrates this for mock observations of the same Earth-like
planet as in Figure~\ref{fig:inference}, but this time taken
over ten days near quadrature, when the disk of the planet is seen at
half phase. The top rows show the thermal and reflected phase
curves and the inferred maps, which look similar to those in
the previous figure.
At the bottom we show the same light curve, but this time including
seven equatorial occultations by a moon one-quarter the size of the planet.
The moon's period and occultation duration are unrealistically short, but
the inferred maps at the bottom show the exquisite constraining power of these
occultations.
In the thermal case, the presence of the moon allows us to infer the
two-dimensional distribution of surface features along its equatorial
occultation path; however, there is little information
in the light curve about features at higher latitudes, and the \starry model
overfits. Conversely, in the reflected light case, there is information
about all latitudes and longitudes, and the inferred map is
largely faithful to the true map.

It should be kept in mind that the kinds of observations
mentioned above will be very challenging for exoplanets, even with
next-generation observatories such as HabEx or LUVOIR. Observations of
occultations of planets by moons, in particular, are likely several decades
away at least. There is some hope that planet-planet occultations may
be detectable in the near future for specific planetary
systems such as TRAPPIST-1
\citep{Luger2019b}, but at extremely limited signal-to-noise.
Even phase curve observations will be difficult
because of their limited signal-to-noise, and in practice many
degeneracies will likely remain. Several studies have found
that color information can greatly help in the interpretation of
reflected phase curves \xxx{(cite)}, while others
have explored in detail the best kinds of priors to assume
\citep[e.g.,][]{Aizawa2020}. Future maps of exoplanets---particularly
terrestrial ones in the habitable zone---will require every tool
in the toolbox.
